PREPRINT -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- PREPRINT.

### 2024-03-12 docs/_tex/index.tex
Laura's comments

\begin{abstract}
Coastal areas support seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
seagrass decline and habitat fragmentation. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the use of satellite data within monitoring programs. Yet,
satellite remote sensing struggles with spatial and spectral resolution,
making it difficult to discriminate seagrass from other macrophytes in
highly heterogenous meadows. To address these challenges and reduce the
issue of mlixed pixel, drone images at a very high spatial resolution
offer a promising solution. This study focuses on using drone
acquisitions with a ten spectral band sensor mirroring those of
Sentinel-2, for mapping intertidal macrophytes and effectively
discriminating between seagrass and green macroalgae. Nine drone flights
were conducted at two different altitudes (12m and 120m) across
heterogeneous European habitats in France and Portugal. Low altitude
flights were used to train a Deep Learning classifier based on Neural
Networks to discrimintate among five taxonomic classes of intertidal
vegetation: Magnoliopsida (Seagrass), Chlorophyceae (Green macroalgae),
Phaeophyceae (Brown algae), Rhodophyceae (Red macroalgae) and
Bacillariophyceae (Diatoms). Classification of drone imagery resulted in
an overall accuracy of 94\% across all the sites and images, covering a
total area of 467 000 m². The model exhibited an accuracy of 96.4\% in
identifying seagrass. This work showed that although they share the same
pigmentary composition, seagrass and green algae can be discriminated
using ten bands multispectral sensor. The algorithm should now be
adapted for Sentinel-2 to see if this discrimination is still
operational at a coarser spatial resolution.
\end{abstract}

@@

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between the land and oceans \citep{unsworth2022}. Seagrass meadows
provide a myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production, protection against sea-level rise and
coastline erosion, and limitation of eutrophication. They serve as vital
habitats for a diverse array of marine and terrestrial species,
providing living, breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation are currently observed due to
climate change, diseases, urbanization, land reclamation, dredging,
competition with alien species, and reduction in water quality
\citetext{\citealp[ ]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp[ ]{duffy2019}; \citealp[
]{rasheed2011long}; \citealp{chefaoui2018dramatic}}. While improvements
in water quality have been recently reported in Europe, allowing an
overall recovery of seagrass ecosystems at local scale, many coastal
waters worldwide are still subjected to strong eutrophication processes
\citetext{\citealp[ ]{deSantos2019}; \citealp{Zoffoli2021}}. Coastal
eutrophication has been ass ociated to excessive accumulation of green
macroalgae, so-called green tides (\citep{devlin2023nutrients}). Green
tides produce shade and suffocation over seagrass individuals, thus
threatening the health of seagrass ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs monitoring Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Traditionally, indicators of seagrass status have been quantified using
\emph{in situ} measurements. The acquisition of field data in intertidal
zones is however notoriously challenging. Intertidal seagrass meadows
are only partially exposed during low tide and can be situated in
difficult-to-reach mudflats, potentially leading to inaccurate and
limited estimations with conventional sampling techniques
\citep{nijland2019}. Satellite observations have been proven effective
in complementing in situ sampling, allowing for the near real-time and
consistent retrieval of seagrass EOVs and EBVs over extensive meadows.
\citetext{\citealp[ ]{Zoffoli2021}; \citealp[ ]{xu2021}; \citealp[
]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing (RS) provides temporally consistent
observations over large spatial scales, it is limited by several
constraints. Satellite missions with a high temporal resolution
(e.g.~daily MODIS observation) are limited by a coarse spatial
resolution (\textgreater100m). Missions with a high spatial resolution
such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low
spectral resolution. The limited number of spectral bands is indeed a
challenge to accurately discriminate seagrass from others co-existing
macrophytes. In particular Chlorophyceae (green algae) and marine
Magnoliopsida (seagrass) share the same pigment composition
\citetext{\citealp[ ]{ralph2002}; \citealp{Douay2022}}. Therefore, to
someone not specialized in the field, their spectral signatures may
appear to be alike \citetext{\citealp[
]{Davies2023}; \citealp{bannari2022}}. Recently, using advanced
machine-learning algorithms trained with a large hyperspectral library
of more then 300 field reflectance spectra, \citep{Davies2023}
demonstrated that it was possible to discriminate Magnoliopsida from
Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral
resolution. However the application of this approach to satellite RS
remains to be validated. Moreover patches of green algae can develop at
small spatial scales that are not observable using non-commercial
satellite imagery \citep{tuya2013}, especially during the initial stage
of th

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in the Gulf of Morbihan (Figure~\ref{fig-map} A), France
(47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only
connected to the sea through a 900m wide channel. A total of 53 small
islands are scattered across the gulf leading to 250 km of shorelines.
Patchy seagrass meadows can be found on a lot of these islands. Within
the Gulf of Morbihan we chose two sites : one located on one of the
gulf's islands (Arz island) and the other located in the southern part
of the gulf (Duer). Two other sites were located in Bourgneuf Bay,
France (46.9849°N, 2.1488°W). This bay is 340 km² semi-enclosed
macrotidal bay, protected from waves by Noirmoutier Island. Bougneuf bay
hosts a large intertidal seagrass meadow of about 6 km². Within this
meadow, the sites observed by drones (L'Epine and Barbatre,
Figure~\ref{fig-map} B) contained a monospecific bed of
\emph{Nanozostera noltei} (dwarf eelgrass) with very little mixing with
other macrophytes. Three sites were surveyed in the Ria de Aveiro Lagoon
in Portugal (40.6887°N, 8.6810°W). The extent of this lagoon is about 80
km² with many narrow channels, large salt marshes and many mudflats that
uncovers at low tide. It is connected to the open sea through a single
channel, with a tidal lag between the North and the South of the lagoon.
The southernmost site (Gafanha) was a mudflat located in a tributary
river of the lagoon whereas the two other sites were situated in the
middle of the lagoon and only accessible by boat (Figure~\ref{fig-map}
C). These Portuguese sites are characterized by a more diverse
intertidal vegetation, where patches of seagrass intermingle with red,
brown and green macroalgae.

@@

\caption{\label{fig-map}Location of drone flights in France and
Portugal. A: Gulf of Morbihan (Two sites), B: Bourngeuf Bay (Two sites),
C: Ria de Aveiro Coastal Lagoon (Three sites). Green represent
intertidal areas.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs with ten spectral bands, from 444
nm (blue) to 840 nm (near infrared, NIR). To ensure consistent lighting
conditions across flight paths, the drone's trajectory was aligned to
maintain a solar azimuth angle of 90 degrees. An overlap of 70\% and
80\% (side and front respectively) between each image was set for each
flight. A downwelling light sensor (DLS2) was used to acquire irradiance
data concomitantly with the camera measurements. Raw data were
calibrated in reflectance using a calibration panel reflective at
\textasciitilde50\% provided by the manufacturer. Across all sites,
flights were made at two different altitudes : 12 m or/and 120 m.
(Table~\ref{tbl-flights}).

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25 m² assessing the presence or absence of five key taxonomic
classes of intertidal vegetation : Bacillariophyceae (Unicellular
benthic diatoms forming biofilms at the sediment surface during low
tide), Phaeophyceae (brown macroalgae), Magnoliopsida (seagrass),
Chlorophyceae (green macroalgae) and Rhodophyceae (red macroalgae)
(Figure~\ref{fig-vegetation}). Pictures of each quadrat were uploaded
online to the Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}, a public open portal to store and share biodiversity
data. Each photograph was also processed to estimate the percent cover
of each type of vegetation using an image processing software (imageJ).
For all quadrats, the hyperspectral reflectance signatures of each
vegetation class was measured using an ASD FieldSpec HandHeld 2
spectroradiometer, which acquires reflectance between 325 and 1075 nm,
with 1 nm of spectral resolution. Hyperspectral signatures from quadrats
serve dual purposes: they validate the radiometric calibration of drone
data and contribute to error reduction in photo interpretation. The
workflow developed in this study is presented in
Figure~\ref{fig-workflow}.

@@

A structure-from-motion photogrammetry software (Agisoft Metashape) was
used to process images to obtain multispectral orthomosaics of each
flight. The workflow for orthomosaicking was the same for every flight.
First, tying key points were detected inside of each image and between
overlapping images in order to obtain a sparse point cloud. This cloud
was cleaned using reprojection accuracy metric in order to remove noisy
points. A dense point cloud was then produced using a structure from
motion algorithm. A surface interpolation of this dense point cloud was
made to obtain a digital surface model (DSM), used to reconstruct the
multispectral ortho-image. Low altitude drone flights produce
ortho-images with a very high spatial resolution (8 mm per pixel),
making it efficient to visually distinguish between the various types of
vegetation. High altitude fights on the other hand allow to cover large
areas and produced images with a pixel size of 80 mm
(Table~\ref{tbl-flights}).

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model with 2 hidden layers. The training
pixels were categorized into seven different classes, representing the
various habitats encountered at the different study sites: Sediment,
Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table~\ref{tbl-flights}) were
used for training because their 8 mm spatial resolution allowied to
avoid spectral sub-pixel mixing and to accurately differentiate
vegetation classes. More than 418,000 pixels at 8 mm resolution from the
3 training flights were used to train the model
(Table~\ref{tbl-validationPX}). Twenty one variables were used by the
model as predictors: the ten raw spectral bands of the Micasense RedEdge
Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same
ten spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{Cao2017}) and the Normalized difference
vegetation index (NDVI, Equation~\ref{eq-ndvi}). Standardisation of
spectral bands is used to eliminate the scaling differences between
spectra and to limit the effect of biomass on the shape of the spectra
\citetext{\citealp[ ]{Douay2022}; \citealp{Davies2023}}.

@@

The model was applied to all flights at both 12 and 120 m of altitude.
In situ information on georeferenced class type and percent cover
collected during each flight was used to assess the model accuracy,
including. These images were used to construct a validation dataset
indicating the presence or absence of each class. Additionally to the
quadrat-based validation dataset, polygons of each class were photo
interpreted in order to increase the number of pixels of the validation
dataset. A confusion matrix, along with precision metrics such as global
accuracy, sensitivity, specificity, and Kappa coefficient, was generated
for each sites. All validation matrices were then merged to create a
unique matrix . Altogether, a total of 536,000 pixels was used to
validate the model, thus providing a geographically robust validation
dataset.

@@

A total of nine prediction maps corresponding to the nine drone flights
were obtained. Each prediction map is associated with a probability map,
indicating the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the five
vegetation classes on which the model was trained, four were present on
this site, with Chlorophyceae and Rhodophyceae mixed with the seagrass
meadow. There was also Bacillariophyceae forming biofilms on parts of
sediment surface. Although the seagrass bed was solely composed of
\emph{Nanozostera noltei}, various colors of this could be observed from
dark green (corresponding to healthy leaves) to whitish/brown (when
leaves were discolored having lost their pigmentation). Regardless of
color, the meadow was accurately predicted as Magnoliopsida by the
model.

@@

The high-altitude flight over Gafanha covered nearly 1 km² in total
(Figure~\ref{fig-GafHigh}). A channel delineating a small island was
masked in the prediction map. Most of the intertidal area was classified
as Magnoliopsida by the model, including seagrass patches with
discolored leaves. Only a few pixels were classified as Chlorophyceae at
this scale. Furthermore, the area that was classified as
Bacillariophyceae in the low-altitude flight remained mostly classified
as such in the high-altitude flight, though some pixels were classified
as Magnoliopsida. Patches of Rhodophyceae were correctly classified. In
the nothern part of the scene, near the land limit, patches of the
schorre angiosperm \emph{Spartina sp.} were misclassified, either as
Magnoliopsida or as Phaeophyceae.

@@

The high altitude flight acquired over the inner lagoon of the Ria de
Aveiro was the largest of all flights, covering almost 1.5 km²
(Figure~\ref{fig-Boat}). The site was dominated by seagrass and the red
macroalgae \emph{Gracilaria vermicullophyla}. The classification
provided consistent results, with a patchy seagrass meadow mixed with
red macroalgae on the eastern part of the scene. As shown in the zoom
(Figure~\ref{fig-Boat}), the edges of the meadow were colonised by green
macroalgae

@@

\caption{\label{fig-Boat}RGB orthomosaic (Top) and Prediction (Bottom)
of the flight made in the inner part of Ria de Aveiro Lagoon, Portugal.
The total extent of this flight was about 1.5 km² with a resolution of
80 mm per pixel. Background colors indicates intertidal area (Light
Green), land area (Light Grey) and water (Light Blue). Each cover an
area equivalent to a 10-meter Sentinel-2 pixel size.}

@@

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike}) was conducted near a dike crossing the northern
part of the scene from west to east. Alongside this dike, brown algae
for the order Fucales (\emph{Fucus spp.}, \emph{Ascophyllum nodosum})
were attached to sparse rocks and stranded green algae (\emph{ulva
spp.}) could be observed Despite the high mixture between Chlorophyceae
and Magnoliopsida these two classes were correctly discriminated by the
classifier.

@@

In this study, a key innovation lies in the utilization of two different
altitudes (12m and 120m) for constructing the neural network model. The
lower altitude flights (8mm spatial resolution) enabled precise
selection of pure pixels representing the classes used in the neural
network model. This methodology implies a consistency between the
reflectance of both altitude. Figure~\ref{fig-CompareRef} depicts the
relationship between reflectance from a low-altitude flight and a
high-altitude flight conducted over the same area. Overall there was a
good agreement between the two altitudes. There was however a slight
underestimation for raw reflectance values in the high-altitude flight,
particularly for higher values (Figure~\ref{fig-CompareRef} A). Since
both flights were conducted over vegetation areas, the highest
reflectance values correspond to the infrared part of the spectrum. This
different is not present anymore when reflectance values of both flights
have been standardized (Equation~\ref{eq-std} ;
Figure~\ref{fig-CompareRef} B).

@@

The computation of the variable importance made it possible to identify
which bands were the most useful for class prediction
(Figure~\ref{fig-VIP}). The spectral bands at 444, 717 and 842 nm of the
Micasense camera did not came out as important for any of the vegetation
classes. The band at 531 nm was the only important predictor for the
classifier to accurately predict Chlorophyceae. In fact, at this
wavelength, the Chlorophyceae spectra have the highest reflectance among
of all vegetation classes. The bands at 531 and 740 nm were the most
important predictors for Phaeophyceae, corresponding to the lowest
reflectance among all classes. Bands at 475 and 560 nm are the most
important predictors for Bacillariophyceae and Rhodophyceae,
respectively. Four predictors, ranging from the green (560 nm) to the
RedEdge (705 nm) bands were important to accurately predict
Magnoliopsida.

@@

\caption{\label{fig-Pigm}Photosynthetic and carotenoid pigments present
(Green) or absent (Red) in each taxonomic class dected by the Neural
Network Classifier, along with their absorption wavelength measured with
spectroradiometer. Chl b: chlorophyll b, Chl c: chlorophyll c, Fuco:
fucoxanthin, Zea: zeaxanthin, Diato: diatoxanthin, Diadino:
diadinoxanthin, Neo: neoxanthin., PE: phycoerythrin, PC: phycocyanin.}

>>>>

\begin{abstract}
Coastal areas support seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
seagrass decline and habitat fragmentation. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the use of satellite data within monitoring programs. Yet,
satellite remote sensing struggles with spatial and spectral resolution,
making it difficult to discriminate seagrass from other macrophytes in
highly heterogenous meadows. Drone images at a very high spatial
resolution offer a promising solution to address challenges related to
spatial heterogeneity and intrapixel mixture. This study focuses on
using drone acquisitions with a ten spectral band sensor mirroring those
of Sentinel-2, for mapping intertidal macrophytes and effectively
discriminating between seagrass and green macroalgae. Nine drone flights
were conducted at two different altitudes (12m and 120m) across
heterogeneous intertidal European habitats in France and Portugal. Low
altitude flights were used to train a Deep Learning classifier based on
Neural Networks to discrimintate among five taxonomic classes of
intertidal vegetation: Magnoliopsida (Seagrass), Chlorophyceae (Green
macroalgae), Phaeophyceae (Brown algae), Rhodophyceae (Red macroalgae)
and benthic Bacillariophyceae (Diatoms). Classification of drone imagery
resulted in an overall accuracy of 94\% across all the sites and images,
covering a total area of 467 000 m². The model exhibited an accuracy of
96.4\% in identifying seagrass. This work showed that although they
share the same pigmentary composition, seagrass and green algae can be
discriminated using a multispectral sensor with ten spectral bands at
444, 475, 531, 560, 650, 668, 705, 717, 740 and 840 nm. The algorithm
should now be adapted for Sentinel-2 to see if this discrimination is
still operational at a coarser spatial resolution.
\end{abstract}

@@

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between the land and oceans \citep{unsworth2022}. Seagrass meadows
provide a myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production, protection against sea-level rise and
coastline erosion, and limitation of eutrophication. They serve as vital
habitats for a diverse array of marine and terrestrial species,
providing living, breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation are currently observed due to
climate change, diseases, urbanization, land reclamation, dredging,
competition with alien species, and reduction in water quality
\citetext{\citealp[ ]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp[ ]{duffy2019}; \citealp[
]{rasheed2011long}; \citealp{chefaoui2018dramatic}}. Both habitat
fragmentation and reduction, in turn, can severely copromise the
effectiveness of ecosystem services provided by seagrass meadows. While
improvements in water quality have been recently reported in Europe,
allowing an overall recovery of seagrass ecosystems at local scale, many
coastal waters worldwide are still subjected to strong eutrophication
processes \citetext{\citealp[ ]{deSantos2019}; \citealp{Zoffoli2021}}.
Coastal eutrophication has been ass ociated to excessive accumulation of
green macroalgae, so-called green tides (\citep{devlin2023nutrients}).
Green tides produce shade and suffocation over seagrass individuals,
thus threatening the health of seagrass ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs to monitor Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Traditionally, indicators of seagrass status have been quantified using
\emph{in situ} measurements. The acquisition of field data in intertidal
zones is however notoriously challenging. Intertidal seagrass meadows
are only partially exposed during low tide and can be situated in
difficult-to-reach mudflats, potentially leading to inaccurate and
limited estimations with conventional sampling techniques
\citep{nijland2019}. Satellite observations have been proven effective
in complementing in situ sampling, allowing for the near real-time and
consistent retrieval of seagrass EOVs and EBVs over extensive meadows.
\citetext{\citealp[ ]{Zoffoli2021}; \citealp[ ]{xu2021}; \citealp[
]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing (RS) provides temporally consistent
observations over large spatial scales, its utilization over intertidal
areas is limited by several constraints. Satellite missions with a high
temporal resolution (e.g.~daily MODIS observation) are limited by a
coarse spatial resolution (\textgreater100m) over these restricted
extension areas. Missions with a high spatial resolution such as
Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low spectral
resolution. The limited number of spectral bands is indeed a challenge
to accurately discriminate seagrass from others co-existing macrophytes.
In particular Chlorophyceae (green algae) and marine Magnoliopsida
(seagrass) share the same pigment composition \citetext{\citealp[
]{ralph2002}; \citealp{Douay2022}}. Therefore, to someone not
specialized in the field, their spectral signatures may appear to be
alike \citetext{\citealp[ ]{Davies2023}; \citealp{bannari2022}}.
Recently, using advanced machine-learning algorithms trained with a
large hyperspectral library of more than 300 field reflectance spectra,
\citep{Davies2023} demonstrated that it was possible to discriminate
Magnoliopsida from Chlorophyceae using radiometric data acquired at
Sentinel-2 's spectral resolution. However the application of this
approach to satellite RS remains to be validated. Moreover patches of
green algae can develop at small spatial scales that are not observable
using non-commercial satellite imagery \citep{tuya2013}, especially
during the initial stage of a green tide.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in the Gulf of Morbihan (Figure~\ref{fig-map} A), France
(47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only
connected to the sea through a 900m wide channel. A total of 53 small
islands are scattered across the gulf leading to 250 km of shorelines.
Patchy seagrass meadows can be found on a lot of these islands. Within
the Gulf of Morbihan we chose two sites : one located on one of the
gulf's islands (Arz island) and the other located in the southern part
of the gulf (Duer). Two other sites were located in Bourgneuf Bay,
France (46.9849°N, 2.1488°W). This bay is 340 km² semi-enclosed
macrotidal bay, protected from waves by Noirmoutier Island. Bourgneuf
bay hosts a large intertidal seagrass meadow of about 6 km². Within this
meadow, the sites observed by drones (L'Epine and Barbatre,
Figure~\ref{fig-map} B) contained monospecific beds of \emph{Nanozostera
noltei} (dwarf eelgrass) with very little mixing with other macrophytes.
Three sites were surveyed in the Ria de Aveiro Lagoon in Portugal
(40.6887°N, 8.6810°W). The extent of this lagoon is about 80 km² with
many narrow channels, large salt marshes and many mudflats that uncovers
at low tide. It is connected to the open sea through a single channel,
with a tidal lag between the North and the South of the lagoon. The
southernmost site (Gafanha) was a mudflat located in a tributary river
of the lagoon whereas the two other sites (Mataducos and Marinha
Lanzarote) were situated in the middle of the lagoon and only accessible
by boat (Figure~\ref{fig-map} C). These Portuguese sites are
characterized by a more diverse intertidal vegetation, where patches of
seagrass intermingle with red, brown and green macroalgae.

@@

\caption{\label{fig-map}Location of drone flights in France and
Portugal. A: Gulf of Morbihan (Two sites), B: Bourngeuf Bay (Two sites),
C: Ria de Aveiro Coastal Lagoon (Three sites). Green represents
intertidal areas.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs with ten spectral bands ranging
from the blue to the near infrared (NIR): 444, 475, 531, 560, 650, 668,
705, 717, 740 and 840 nm. To ensure consistent lighting conditions
across flight paths, the drone's trajectory was aligned to maintain a
solar azimuth angle of 90 degrees. An overlap of 70\% and 80\% (side and
front respectively) between each image was set for each flight. A
downwelling light sensor (DLS2) was used to acquire irradiance data
concomitantly with the camera measurements. Raw data were calibrated in
reflectance using a calibration panel reflective at \textasciitilde50\%
provided by the manufacturer. Across all sites, flights were made at two
different altitudes : 12 m or/and 120 m. (Table~\ref{tbl-flights}).

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25 m² assessing the presence or absence of five key taxonomic
classes of intertidal vegetation : Bacillariophyceae (Unicellular
benthic diatoms forming biofilms at the sediment surface during low
tide), Phaeophyceae (brown macroalgae), Magnoliopsida (dwarf seagrass),
Chlorophyceae (green macroalgae) and Rhodophyceae (red macroalgae)
(Figure~\ref{fig-vegetation}). Pictures of each quadrat were uploaded
online to the Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}, a public open portal to store and share biodiversity
data. Each photograph was also processed to estimate the percent cover
of each type of vegetation using an image processing software (imageJ).
For all quadrats, the hyperspectral reflectance signatures of each
vegetation class was measured using an ASD FieldSpec HandHeld 2
spectroradiometer, which acquires reflectance between 325 and 1075 nm,
with 1 nm of spectral resolution. Hyperspectral signatures from quadrats
serve for dual purposes: they validate the radiometric calibration of
drone data and contribute to error reduction in photo interpretation.
The workflow developed in this study is presented in
Figure~\ref{fig-workflow}.

@@

A structure-from-motion photogrammetry software (Agisoft Metashape) was
used to process images to obtain multispectral orthomosaics of each
flight. The workflow for orthomosaicking was the same for every flight.
First, tying key points were detected inside of each image and between
overlapping images in order to obtain a sparse point cloud. This cloud
was cleaned using reprojection accuracy metric in order to remove noisy
points. A dense point cloud was then produced using a structure from
motion algorithm. A surface interpolation of this dense point cloud was
made to obtain a digital surface model (DSM), used to reconstruct the
multispectral ortho-image. Low altitude drone flights produced
ortho-images with a very high spatial resolution (8 mm per pixel),
making it efficient to visually distinguish between the various types of
vegetation. High altitude fights on the other hand allowed to cover
large areas and produced images with a pixel size of 80 mm
(Table~\ref{tbl-flights}).

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model with 2 hidden layers. The training
pixels were categorized into seven different classes, representing the
various habitats encountered at the different study sites: Sediment,
Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table~\ref{tbl-flights}) were
used for training because their 8 mm spatial resolution allowied to
avoid spectral sub-pixel mixing and to accurately identify vegetation
classes. More than 418,000 pixels at 8 mm resolution from the 3 training
flights were used to train the model (Table~\ref{tbl-validationPX}).
Twenty one variables were used by the model as predictors: the ten raw
spectral bands of the Micasense RedEdge Dual MX multispectral camera
(ranging from 444 nm to 840 nm), the same ten spectral bands
standardized using a min/max transformation (Equation~\ref{eq-std} ;
\citep{Cao2017}) and the Normalized difference vegetation index (NDVI,
Equation~\ref{eq-ndvi}). Standardisation of spectral bands is commonly
used to eliminate the scaling differences between spectra and to limit
the effect of biomass on the spectra shape \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

@@

The model was applied to all flights at both 12 and 120 m of altitude.
In situ information on georeferenced class type and percent cover
collected during each flight was used to assess the model accuracy.
These images were used to construct a validation dataset indicating the
presence or absence of each class. Additionally to the quadrat-based
validation dataset, polygons of each class were photo interpreted in
order to increase the number of pixels of the validation dataset. A
confusion matrix, along with precision metrics such as global accuracy,
sensitivity, specificity, and Kappa coefficient, was generated for each
sites. All validation matrices were then aggregated to create a unique
matrix . Altogether, a total of 536,000 pixels was used to validate the
model, thus providing a geographically robust validation dataset.

@@

The nine drone flights were utilized to produce a total of nine
prediction maps, each of them associated with a probability map that
indicates the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the five
vegetation classes on which the model was trained, four were present on
this site, with Chlorophyceae and Rhodophyceae mixed with the seagrass
meadow. There was also Bacillariophyceae forming biofilms on parts of
sediment surface. Although the seagrass bed was solely composed of
\emph{Nanozostera noltei}, various colors of this specie could be
observed from dark green (corresponding to healthy leaves) to
whitish/brown (when leaves were discolored having lost their
pigmentation). Regardless of color, Magnoliopsida pixels were accurately
predicted by the model.

@@

The high-altitude flight over Gafanha covered a total area of
approwimately 1 km² (Figure~\ref{fig-GafHigh}). A channel contouring a
small island was masked in the prediction map. Most of the intertidal
area was classified as Magnoliopsida by the model, including seagrass
patches with discolored leaves. Only a few pixels were classified as
Chlorophyceae at this spatial scale. Furthermore, the area that was
classified as Bacillariophyceae in the low-altitude flight remained
mostly classified as such in the high-altitude flight, though some
pixels were classified as Magnoliopsida. Patches of Rhodophyceae were
correctly classified. In the northern part of the site and near the land
eadges, patches of the schorre angiosperm \emph{Spartina sp.} were
misclassified, either as Magnoliopsida or as Phaeophyceae.

@@

Amoung the high altitude flights, the one acquired over the inner lagoon
of the Ria de Aveiro covered the largest area with approximately 1.5 km²
(Figure~\ref{fig-Boat}). This site was dominated by seagrass and the red
macroalgae \emph{Gracilaria vermicullophyla}. The classification
provided consistent results, with a patchy seagrass meadow mixed with
red macroalgae on the eastern part of the site. As shown in the zoom
(Figure~\ref{fig-Boat}), the edges of the meadow were colonised by green
macroalgae

@@

\caption{\label{fig-Boat}RGB orthomosaic (Top) and Prediction (Bottom)
of the flight made in the inner part of Ria de Aveiro Lagoon, Portugal.
The total extent of this flight was about 1.5 km² with a resolution of
80 mm per pixel. Background colors indicate intertidal area (Light
Green), land area (Light Grey) and water (Light Blue). Each cover an
area equivalent to a 10 m Sentinel-2 pixel size.}

@@

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike}) was conducted near a dike, which crossed the
northern part of the site from west to east. Alongside this dike, brown
algae for the order Fucales (\emph{Fucus spp.}, \emph{Ascophyllum
nodosum}) were attached to sparse rocks and stranded green algae
(\emph{Ulva spp.}) could be observed Despite the high mixture between
Chlorophyceae and Magnoliopsida these two classes were correctly
discriminated by the classifier.

@@

In this study, a key innovation lies in the utilization of drone flights
at two different altitudes (12m and 120m) for constructing the neural
network model. The lower altitude flights (8mm spatial resolution)
enabled precise selection of pure pixels to train the neural network
model. This methodology implies a consistency between the reflectance of
both altitudes. Figure~\ref{fig-CompareRef} depicts the relationship
between reflectance from a low and a high-altitude flights conducted
over the same area. Overall there was a good agreement between the two
altitudes. There was however a slight underestimation for raw
reflectance values in the high-altitude flight, particularly for higher
reflectance values (Figure~\ref{fig-CompareRef} A). Since both flights
were conducted over vegetation areas, the highest reflectance values
correspond to the infrared part of the spectrum. This different is not
present anymore when reflectance values of both flights have been
standardized (Equation~\ref{eq-std} ; Figure~\ref{fig-CompareRef} B).

@@

The computation of the variable importance made it possible to identify
which bands were the most useful for class prediction
(Figure~\ref{fig-VIP}). The spectral bands at 444, 717 and 842 nm of the
Micasense camera did not provide as important information to
discriminate any of the vegetation classes. The band at 531 nm was the
only important predictor for the classifier to accurately predict
Chlorophyceae. In fact, at this wavelength, the Chlorophyceae spectra
showed the highest reflectance among of all vegetation classes. The
bands at 531 and 740 nm were the most important predictors for
Phaeophyceae, corresponding to the lowest reflectance among all classes.
Bands at 475 and 560 nm were the most important predictors for
Bacillariophyceae and Rhodophyceae, respectively. Four predictors,
ranging from the green (560 nm) to the RedEdge (705 nm) bands were
important to accurately predict Magnoliopsida.

@@

\caption{\label{fig-Pigm}Photosynthetic and carotenoid pigments present
(Green) or absent (Red) in each taxonomic class present in the Neural
Network Classifier, along with their absorption wavelength measured with
spectroradiometer. Chl b: chlorophyll b, Chl c: chlorophyll c, Fuco:
fucoxanthin, Zea: zeaxanthin, Diad: diadinoxanthin, Lut: lutein, Neo:
neoxanthin, PE: phycoerythrin, PC: phycocyanin.}
